Supplemental Materials

### 2025-01-10 products/manuscript/Manuscript_Final_SSM.qmd
Final Commit
After copy edits

The differences between and within groups both today and in the past have prompted racially-stratified research. While race is a social construct and thus biological fiction, the social consequences of racism influence health and disease patterns, including birth outcomes, particularly in countries with high levels of race-based inequality [@meloni2022; @travassos2004]. Contemporary studies in Brazil have shown that non-White mothers have higher odds of giving birth to LBW infants [@nyarko2013; @wehby2015a; @vilanova2019]. Non-White mothers also have infants with lower mean birth weights than their White counterparts [@leal2006]. These trends also hold true in the United States, the other major slave society in the Americas, where today Black newborns are at higher risk for LBW than White newborns, with trends worsening over time [@pollock2021]. Racism is a possible upstream contributor, as studies demonstrate that U.S.-born Black women have higher risk for both LBW and PTB than Black immigrant women [@braveman2021a; @david1997].

@@

The journal is held at the Biblioteca Nacional (BN), the Maternidade Escola, Rio de Janeiro (ME-UFRJ), and the Biblioteca CB/A-Biomédica A, Universidade Estadual do Rio de Janeiro (UERJ), all in Rio de Janeiro, Brazil. Between January 2012 and July 2013, I photographed each volume. Then, between January and August 2017, I manually input the data into Excel from the digital reproductions. I then converted this into a .csv file for upload into R version 4.4.1 (2024-06-14).

@@

*Maternal parity*. The original hospital records divided maternal parity and gravidity into eight categories of parity and gravidity: nulliparous, primiparous, secundiparous, and multiparous; and primigravida, secundigravida, trigravida, and multigravida. Due to issues of statistical power, I combined these into two categories: 1) `nulliparous`, which included nulliparous, primigravida, and primiparous (any woman giving birth for the first time); and 2) `multiparous`, which included secundiparous, multiparous, secundigravida, trigravida, and multigravida (any woman who has given birth more than once).

@@

*Year*. I construct year fixed effects by including dummy variables for each year from 1923 to 1926. 1922 is the base year. The coefficient on these dummy variables is the year fixed effect, which indicates how the average birth weight is different from the base year.

@@

Table 1 provides summary statistics for the dataset. Of hospital patients, 58% were of African descent (defined as *preta* or *parda*) and 42% were of European descent (*branca*). An average of census data for the city's female population's racial makeup between 1890 and 1940, years for which data exist, show that White women comprised 63%, mixed-race women 24%, and Black women 14% (with rounding errors) [@roth2020]. Thus, at Laranjeiras, women of color were over-represented among hospital patients. First-time mothers (nulliparous), comprised 41%. Of all women, 84% were Brazilian. Of the nearly 16% immigrant patients, 15% were European.

@@

In all models (Tables 2 and 3), women who were nulliparous (delivering their first child) had lighter infants weighing than multiparous women, an association recognized by physicians at the time and found in both other historical studies and today [@conrado1924a; @costa1998; @souza2020]. Age and nationality were not significant. The year fixed effects variable demonstrates that birthweights were highest in 1922.

@@

As scholars have shown, mean birth weights in the past were not drastically lower than today, and Laranjeiras fits this pattern, with birth weights relatively close to those of contemporary Brazilian infants even if lower than infants born in other regions during the same period [@costa1998; @costa2004; @goldin1989]. Across populations, relatively high mean birth weights in the nineteenth and early twentieth centuries could be an artifact of high stillbirth rates, which selected upon fetuses that were already small and unhealthy [@costa1998; @costa2004; @schneider2017]. Today, PTB and lower birth weight infants that might have died in utero 100 years ago now survive, bringing the mean birth weight down.

@@

Two sets of missing data are notable. First, hospital records excluded maternal marital status (included in 0.4% of records). Scholars have found that marital status was a good proxy for SES [@costa2004; @goldin1989]. The hospital's exclusion of this variable precludes a full understanding of mothers' SES.

@@

This study has shown that lower birth weights were associated with Afro-descended women giving birth in the Maternidade Laranjeiras in Rio de Janeiro, Brazil between 1922 and 1926, suggesting that racial inequality stemming from slavery impacted health conditions in the post-abolition period, possibly through multiple mechanisms for which we do not have direct evidence. These findings, although not generalizable beyond this specific hospital, are consistent with both historical and present-day studies on racialized disparities in birth weight in Rio de Janeiro, demonstrating how racial inequality in health outcomes has persisted over time.

>>>>

The differences between and within groups both today and in the past have prompted research stratified by race. While race is a social construct and thus biological fiction, the social consequences of racism influence health and disease patterns, including birth outcomes, particularly in countries with high levels of race-based inequality [@meloni2022; @travassos2004]. Contemporary studies in Brazil have shown that non-White mothers have higher odds of giving birth to LBW infants [@nyarko2013; @wehby2015a; @vilanova2019]. Non-White mothers also have infants with lower mean birth weights than their White counterparts [@leal2006]. These trends also hold true in the United States, the other major slave society in the Americas, where today Black newborns are at higher risk for LBW than White newborns, with trends worsening over time [@pollock2021]. Racism is a possible upstream contributor, as studies demonstrate that U.S.-born Black women have higher risk for both LBW and PTB than Black immigrant women [@braveman2021a; @david1997].

@@

The journal is held at the Biblioteca Nacional (BN); the Maternidade Escola, Rio de Janeiro (ME-UFRJ); and the Biblioteca CB/A-Biomédica A, Universidade Estadual do Rio de Janeiro (UERJ), all in Rio de Janeiro, Brazil. Between January 2012 and July 2013, I photographed each volume. Then, between January and August 2017, I manually input the data into Excel from the digital reproductions. I then converted this into a .csv file for upload into R version 4.4.1 (2024-06-14).

@@

*Maternal parity*. The original hospital records divided maternal parity and gravidity into eight categories of parity and gravidity: nulliparous, primiparous, secundiparous, and multiparous; and primigravida, secundigravida, trigravida, and multigravida. Due to issues of statistical power, I combined these into two categories: 1) `nulliparous`, including nulliparous, primigravida, and primiparous (any woman giving birth for the first time); and 2) `multiparous`, including secundiparous, multiparous, secundigravida, trigravida, and multigravida (any woman who has given birth more than once).

@@

*Year*. I constructed year fixed effects by including dummy variables for each year from 1923 to 1926. 1922 is the base year. The coefficient on these dummy variables is the year fixed effect, which indicates how the average birth weight is different from the base year.

@@

Table 1 provides summary statistics for the dataset. Of hospital patients, 58% were of African descent (defined as *preta* or *parda*) and 42% were of European descent (*branca*). An average of census data for the city's female population's racial makeup between 1890 and 1940, years for which data exist, show that White women comprised 63%, mixed-race women 24%, and Black women 14% (with rounding errors) [@roth2020]. Thus, at Laranjeiras, women of color were over-represented among hospital patients. First-time mothers (nulliparous), comprised 41%. Of all women, 84% were Brazilian, 15% were European immigrants, and the remaining were immigrants from other regions.

@@

In all models (Tables 2 and 3), women who were nulliparous (delivering their first child) had lighter infants than multiparous women, an association recognized by physicians at the time and found in both other historical studies and today [@conrado1924a; @costa1998; @souza2020]. Age and nationality were not significant. The year fixed effects variable demonstrates that birthweights were highest in 1922.

@@

As scholars have shown, mean birth weights in the past were not drastically lower than today, and Laranjeiras fits this pattern, with birth weights relatively close to those of contemporary Brazilian infants even if lower than infants born in other regions during the same period [@costa1998; @costa2004; @goldin1989]. Across populations, relatively high mean birth weights in the nineteenth and early twentieth centuries could be an artifact of high stillbirth rates, which selected upon fetuses that were already small and unhealthy [@costa1998; @costa2004; @schneider2017]. Today, PTB and lower birth weight infants that might have died in utero 100 years ago survive, bringing the mean birth weight down.

@@

Two sets of missing data are notable. First, hospital records excluded maternal marital status (included in only 0.4% of records). As scholars have found that marital status was a good proxy for SES [@costa2004; @goldin1989], the hospital's exclusion of this variable precludes a full understanding of mothers' SES.

@@

This study has shown that lower birth weights were associated with Afro-descended women giving birth in the Maternidade Laranjeiras in the city of Rio de Janeiro, Brazil between 1922 and 1926, suggesting that racial inequality stemming from slavery impacted health conditions in the post-abolition period, possibly through multiple mechanisms for which we do not have direct evidence. These findings, although not generalizable beyond this specific hospital, are consistent with both historical and present-day studies on racialized disparities in birth weight in Rio de Janeiro, demonstrating how racial inequality in health outcomes has persisted over time.

### 2025-01-10 products/manuscript/supplement/Supplemental_Materials.qmd
Final Commit
After copy edits

```{r}
#| label: fig-schematic-4
#| fig-cap: Scatterplot of birth weight by length stratified by sex
#| echo: FALSE
knitr::include_graphics(here("results","figures","length_weight_sex.png"))
```
I also create a plot of the distribution of birth weights at Laranjeiras, following Wilcox's [-@wilcox2001] discussion of the residual and predominant birth weight distributions.

@@

I also calculate sex ratio at birth (SRaB) in the hospital:

@@

The third exploratory statistical model is a multilinear regression, looking at the relationship of maternal skin color, age, age squared, nationality, and gestational status on infant birth weight with the `ModifiedColor` variable as the main predictor variable.

>>>>

```{r}
#| label: fig-schematic-4
#| fig-cap: Scatterplot of birth weight by length stratified by sex
#| echo: FALSE
knitr::include_graphics(here("results","figures","length_weight_sex.png"))
```
I also created a plot of the distribution of birth weights at Laranjeiras, following Wilcox's [-@wilcox2001] discussion of the residual and predominant birth weight distributions.

@@

I also calculated sex ratio at birth (SRaB) in the hospital:

@@

The third exploratory statistical model is a multilinear regression, looking at the relationship of maternal skin color, age, $age^2$, nationality, and gestational status on infant birth weight with the `ModifiedColor` variable as the main predictor variable.
